---
title: "Introduction to R for Biologists"
subtitle: "BIOL 4994/BIOL 4991 undegrads <br> BIOL 6994/BIOL 6997 grads"
author: "Alan J. Figueroa Ruiz"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
output:
  prettydoc::html_pretty:
  theme: cayman
  highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Deadline / Fecha de entrega
**Deadline:** Friday, Nov 31, 2025; 11:59 PM.  
**Fecha de entrega:** Viernes, 31 de noviembre de 2025; 11:59 PM.

### Instructions / Instrucciones generales

- This assignment consists of **uploading a plot and a brief paragraph (~250 words) summarizing your findings**.
(Esta tarea consiste en subir un gráfico y un párrafo breve (~250 palabras) resumiendo tus hallazgos.)

- In your paragraph, include the background of the data, what you did, how you did it, and what you found.
(En tu párrafo, incluye el contexto de los datos, qué hiciste, cómo lo hiciste y qué encontraste.)

- Submit your HTML file exported/knit from the .Rmd. Do not submit screenshots unless requested.
(Entrega el archivo HTML generado/kniteado desde el .Rmd. No envíes capturas de pantalla a menos que se indique.)

- Your paragraph should briefly summarize your workflow and findings. For example:
(Tu párrafo debe resumir brevemente tu flujo de trabajo y hallazgos. Por ejemplo:)

- I analyzed the data of tyre performance. I visualized it using ggplot() and then ran a one-way ANOVA. We checked for normality of residuals and found that the residuals were approximately normally distributed. I ran a Tukey post-hoc test. I found that the high-performance tyres were Brand X, and the lower-performance tyres were Brand Y and Z.

- Analicé los datos del desempeño de los neumáticos. Los visualicé usando ggplot() y luego ejecutamos un ANOVA de una vía. Verificamos la normalidad de los residuos y encontramos que los residuos estaban aproximadamente distribuidos normalmente. Ejecuté la prueba post-hoc de Tukey. Encontré que los neumáticos de alto desempeño fueron la marca X y los de menor desempeño fueron las marcas Y y Z.

<br>

Uploading data and head to ensure csv file was read correctly
```{r}
tyre <- read.csv("tyre.csv")
head(tyre, 20)
```

<br>

Libraries instalation and plot construction
```{r, message = FALSE}
library(ggplot2)
library(car)
library(dplyr)


p <- ggplot(tyre, aes(x=Brands, y=Mileage)) +
  theme_classic() +
  geom_boxplot(aes(fill = Brands)) + # Esteticidad
  labs(title= "Rendimiento de Gomas mas Populares en Puerto Rico", x="Marcas de Gomas",
       y="Millaje (Km)", fill = "Marcas") + #Labels
  theme(plot.title = element_text(size=18, face="bold", hjust = 0.5, color = "red")) + #themes       graph title
  theme(axis.title.y = element_text(size=14, face="bold", color = "blue")) + #themes Y title
  theme(axis.text.y = element_text(size=14, face="bold", color = "black")) + #themes Y text title
  theme(axis.title.x = element_text(size=14, face="bold", color = "blue")) + #themes X title
  theme(axis.text.x = element_text(size=14, face="bold", color = "black")) + #themes Y text title
  theme(legend.title = element_text(size=18, face="plain", color = "red")) #legend text title
p
```
```{r, echo=FALSE}
ggsave(filename = "plot.png",
       plot = p,
       width=8, height = 6, dpi = 300)
```

<br>

Developing some ANOVA statistics on the model to know what particular is happening. We establish on class as first assumption for ANOVA that all residues follow a normal distribution.
```{r}
mod <-aov(Mileage ~ Brands, data = tyre) #Construccion de modelo
summary(mod) #Resultados de prueba ANOVA
resid(mod) #Mostrar residuales
resid_anova <- resid(mod) #Asignar residuales
shapiro.test(resid_anova) #Prueba de normalidad de residuales, Probar segunda asumcion de ANOVA 
suppressWarnings(leveneTest(Mileage ~ Brands, data=tyre)) #Probar  homogenidad de las varianzas, Probar tercera asumpción de ANOVA
TukeyHSD(mod) #Comparacion Multiple de las Medias
```

```{r, echo=FALSE}
ggsave(filename = "plot.png",
       plot = p,
       width=8, height = 6, dpi = 300)
```

Results Summary

Se analizó el conjunto de datos tyre.csv con el objetivo de comparar el desempeño de distintas marcas de neumáticos en términos de kilometraje. Primero, se visualizó los datos utilizando la función ggplot() para observar la distribución del rendimiento entre las marcas. Luego, se construyo un modelo ANOVA de una vía (aov(Mileage ~ Brands, data = tyre)) para evaluar si existían diferencias significativas en el kilometraje promedio según la marca. Los resultados del ANOVA mostraron un efecto significativo de la marca sobre el rendimiento (F = 17.94, p < 0.001), lo que indica que no todas las marcas ofrecen el mismo desempeño. Luego, se verificó los supuestos del modelo. La prueba de Shapiro–Wilk (W = 0.9872, p = 0.7826) confirmó que los residuos seguían una distribución aproximadamente normal, y la prueba de Levene (p = 0.5592) mostró que las varianzas eran homogéneas entre grupos. Cumplidos los supuestos, se aplicó la prueba de Tukey HSD para identificar qué marcas diferían significativamente entre sí. Los resultados revelaron que los neumáticos Falken presentaron un kilometraje significativamente mayor en comparación con Apollo, CEAT y Bridgestone, con valores de p ajustados inferiores a 0.01. En contraste, Bridgestone obtuvo el menor rendimiento promedio.En resumen, el análisis demuestra que la marca influye de manera significativa en el desempeño de los neumáticos por millaje en Puerto Rico, siendo Falken la de mejor rendimiento y Bridgestone la de menor. Este tipo de análisis estadístico permite respaldar decisiones basadas en evidencia en evaluaciones de producto y control de calidad.

The tyre.csv dataset was analyzed with the goal of comparing the performance of different tire brands in terms of mileage. First, the data were visualized using the ggplot() function to observe the distribution of performance among brands. Then, a one-way ANOVA model (aov(Mileage ~ Brands, data = tyre)) was constructed to evaluate whether there were significant differences in average mileage across brands. The ANOVA results showed a significant effect of brand on performance (F = 17.94, p < 0.001), indicating that not all brands provide the same level of performance. Next, the model assumptions were verified. The Shapiro–Wilk test (W = 0.9872, p = 0.7826) confirmed that the residuals were approximately normally distributed, and the Levene’s test (p = 0.5592) showed that variances were homogeneous across groups. Once the assumptions were met, the Tukey HSD test was applied to identify which brands differed significantly from each other. The results revealed that Falken tires had significantly higher mileage compared to Apollo, CEAT, and Bridgestone, with adjusted p-values below 0.01. In contrast, Bridgestone showed the lowest average performance. In summary, the analysis demonstrates that tire brand significantly influences mileage performance in Puerto Rico, with Falken showing the best results and Bridgestone the lowest. This type of statistical analysis supports evidence-based decision-making in product evaluation and quality control.